TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-10 paper/draft.tex
minor intro changes

A number of studies focused on improving M dwarf metallicity precisions have used systems of M dwarfs in common proper motion with an FGK star, to developed precise empirical relations (as good as $\sim$0.07\,dex) between the adopted metallicities of the FGK primary (under the assumption kinematically associated wide binaries are co-evolutionary; \citealt{Bonfils:2005}), and strong, isolated lines in the spectra of the M dwarf (e.g. \citealt{Terrien:2012}, \citealt{Rojas-Ayala:2012}, \citealt{Newton:2014}). However these metallicity calibrations are inaccessible for analysis with \apogee, as the wavelength coverage does not include important calibration indices in the $K$-band or include important information about the spectral shape needed to determine $\teff$. Further, earlier calibrations are generally based on moderate-resolution data \citep[with some exceptions,][]{Neves:2014, Lindgren:2016} that would fail to utilize the additional spectral information provided by \apogee's resolution.

@@

Here we take M-dwarf labels from samples of optically observed stars that overlap the
\sdssiv\ \apogee\ sample, and use those labels to train a model and label all of the
\sdssiv\ \apogee\ M dwarfs.
One set of labels are physical parameters (effective temperatures and metallicities)
and the other set of labels are spectral types.
In each case, \thecannon\ produces a predictive model for \apogee\ spectra, but without
any input of stellar physics beyond whatever went in to producing the labels in the first
place, which was in the context of optical spectroscopy. That is, the model to be applied to the \apogee\ data is learned from the \apogee\ data alone.
This obviates physical models in the infrared, and capitalizes on the excellent data quality
delivered by the \apogee\ spectrograph and project.

>>>>

A number of studies focused on improving M dwarf metallicity precisions have used systems of M dwarfs in common proper motion with an FGK star, to developed precise empirical relations (as good as $\sim$0.07\,dex) between the adopted metallicities of the FGK primary (under the assumption kinematically associated wide binaries are co-evolutionary; \citealt{Bonfils:2005}), and strong, isolated lines in the spectra of the M dwarf (e.g. \citealt{Terrien:2012}, \citealt{Rojas-Ayala:2012}, \citealt{Newton:2014}). However these metallicity calibrations are inaccessible for analysis with \apogee, as the wavelength coverage does not include important calibration indices in the $K$-band and does not contain information about the overall spectral shape often used to determine $\teff$ and spectral type. Further, earlier calibrations are generally based on moderate-resolution data \citep[with some exceptions,][]{Neves:2014, Lindgren:2016} that fail to utilize the additional spectral information provided by \apogee's resolution.

@@

Here we take M-dwarf labels from samples of well-characterized stars that overlap the
\sdssiv\ \apogee\ sample, and use those labels to train a model and label all of the M dwarfs observed by
\sdssiv\ \apogee. One set of labels are physical parameters (effective temperatures and metallicities)
and the other set of labels are spectral types. In each case, \thecannon\ produces a predictive model for \apogee\ spectra, but without
any input of stellar physics beyond whatever went in to producing the labels in the first
place. That is, the model to be applied to the \apogee\ data is learned from the \apogee\ data alone.
This obviates physical models that are often inaccurate for M dwarfs, and capitalizes on the excellent data quality
delivered by the \apogee\ spectrograph and project.
